Supplementary Information

### 2020-11-26 doc/manuscript/MEE_manuscript.Rmd
add text on long-term trends, work on Paolo's feedback

**Tree rings provide a long-term record of annual growth increments that is invaluable for understanding forests in an era of global change.** 
Spanning time scales of decades to centuries or even millennia, they provide by far the most robust method for characterization of the interannual climate sensitivity of tree growth [@fritts_tree_1976] and how it is changing [e.g., @sniderhan_growth_2016; @maxwell_declining_2016]. 
Combined with forest censuses, they can be used to estimate forest woody productivity [$ANPP_{woody}$; @graumlich_long-term_1989; @teets_linking_2018] and its climate sensitivity [@klesse_sampling_2018; @teets_quantifying_2018; @helcoski_growing_2019]. 
This information is critical to predicting forest responses to climate change, and thereby reducing the enormous uncertainty surrounding future contributions of Earth's forests to the global carbon cycle [@friedlingstein_climatecarbon_2006].
Yet, dendrochronological records have been traditionally collected and analyzed to detect climate signals rather than to predict forest productivity and its climate sensitivity [@klesse_sampling_2018].
As a result, prevailing approaches hold a number of limitations for using tree-rings to address pressing questions concerning forest productivity in the current era of rapid environmental change.

@@

**Characterizing how tree growth is responding to global change (slowly-changing environmental drivers) is very challenging and uncertain.**

>>>>

**Tree rings provide a long-term record of annual growth increments that is invaluable for understanding forests in an era of global change.** 
Spanning time scales of decades to centuries or even millennia, they provide by far the most robust method for characterization of the interannual climate sensitivity of tree growth [@fritts_tree_1976] and how it is changing [e.g., @sniderhan_growth_2016; @maxwell_declining_2016]. 
Combined with forest censuses, they can be used to estimate forest woody productivity [$ANPP_{woody}$; @graumlich_long-term_1989; @teets_linking_2018] and its climate sensitivity [@klesse_sampling_2018; @teets_quantifying_2018; @helcoski_growing_2019]. 
*other global change drivers*
This information is critical to predicting forest responses to climate change, and thereby reducing the enormous uncertainty surrounding future contributions of Earth's forests to the global carbon cycle [@friedlingstein_climatecarbon_2006].
Yet, collection and analysis of dendrochronological records has been traditionally been optimized to detect climate signals rather than to predict forest productivity, its climate sensitivity, and how it may be changing [@cherubini_potential_1998; @nehrbassahles_influence_2014; @klesse_sampling_2018].
As a result, prevailing approaches hold a number of limitations for using tree-rings to address pressing questions concerning forest productivity in the current era of rapid environmental change.

@@

**Characterizing how tree growth and forest productivity are responding to global change (slowly-changing environmental drivers) is very challenging and uncertain.**

- directional climate change (as opposed to variation), CO~2~, and atmospheric deposition are all potentially influencing tree growth.

- at the same time, stand dynamics influence growth; many forests are secondary, and ANPP_woody may decline as stands age

- a variety of methods are used to account for simultaneous changes in tree size, including... (DENDRO_REFS).

- One commonly applied approach is to assume that $BAI$ is fundamentally independent of size (DENDRO_REFS).

- None of these methods is fully satisfactory, being subject to various sampling and analysis biases [@cherubini_potential_1998; @nehrbassahles_influence_2014], leaving great uncertainty as to what the tree-ring record can tell us about the impact of rising CO~2~ [@walker_integrating_2020].

- It is necessary to *simultaneously* account for the influences of changing tree size and environmental drivers.
